Rate Transfers with Polyhedral Surrogate Losses /\\ Surrogate Regret Bounds for Polyhedral Losses

### 2021-05-20 rates-neurips-21/rates.tex
upper bound stuff

\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
\usepackage{bbm}
\usepackage{amsmath,amsfonts,amssymb}

@@

\paragraph{Target problems.}
We consider supervised machine learning problems where data points have features in $\X$ and labels in $\Y$, a finite set.
The \emph{target problem} is to learn a hypothesis mapping $\X$ to a \emph{report} (or \emph{prediction}) \emph{space} $\R$, possibly different from $\Y$.
In this paper, we assume $\R$ is a finite set.
%For example, standard binary classification has $\Y = \R = \{-1,+1\}$, whereas in a ranking problem, $\R$ is the set of all permutations of $\Y$.
We suppose the target problem is specified by the \emph{target loss} $\ell:\R\to\reals^\Y_+$, which maps a report (prediction) $r \in \R$ to the vector $\ell(r)$ of nonnegative loss values for each label, i.e. $\ell(r) = (\ell(r)_y)_{y\in\Y}$.
Given a hypothesis $g: \X \to \R$ and a data point $(x,y) \in \X \times \Y$, the loss is $\ell(g(x))_y$.
\bo{We will make a minimal non-redundancy assumption on $\ell$ in [Ref].}

@@

\bo{all in progress}
The following definition applies both to target losses and surrogate losses.
\begin{definition}[Elicits, report space]
  \label{def:elicits}
  A loss function $f:R\to\reals^\Y_+$, where the domain $R$ is called the \emph{report space}, \emph{elicits} a property $\Gamma:\simplex \toto \R$ if
  As each $f$ elicits a unique property, we refer to it as $\prop{f}$.
\end{definition}
In particular, we will consistently use the notation $\Gamma = \prop{L}$ where $L$ is a surrogate loss with report space $\reals^d$, and we will use $\gamma = \prop{\ell}$ where $\ell$ is a target loss with finite report space $\R$.

@@

\paragraph{Calibration and consistency.}
It is known that consistency implies the weaker condition of \emph{calibration}, which implies the still-weaker, but very useful, condition of \emph{indirect elicitation}.
To define these conditions, consider $L,\psi,\ell$ and $\gamma = \prop{\ell}$.
Recall that $\gamma(p)$ of target reports that are optimal for $p$.
If $\psi(u) \not\in \gamma(p)$, then $u$ is a ``bad'' (suboptimal) surrogate prediction.
Let $B_{L,\psi,\ell}(p) = \{\inprod{p}{L(u)} ~:~ \psi(u) \not\in \gamma(p)\}$, the set of expected surrogate losses resulting from bad predictions.
\begin{definition}[Indirectly elicits, Calibrated]
  A surrogate-link pair $(L,\psi)$ \emph{indirectly elicits} a target $\ell$ if $0 \not\in B_{L,\psi,\ell}(p)$ for all $p \in \simplex$.
  Meanwhile, $(L,\psi)$ is \emph{calibrated} for $\ell$ if $0 < \inf B_{L,\psi,\ell}(p)$ for all $p \in \simplex$.
%  \begin{equation}
%    \inf_{u : \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)} > \risk{L}(p)$ .
%  \end{equation}
%  Throughout, $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
\end{definition}
Equivalently, $(L,\psi)$ indirectly elicits $\ell$ if $u \in \Gamma(p) \implies \psi(u) \in \gamma(p)$.
Calibration is the slightly more subtle condition that there is a nonzero gap between the expected loss of a ``good'' report and that of any bad one.
%Observe that indirect elicitation is the condition that $0 \not\in \{\inprod{p}{L(u)} ~:~ \psi(u) \not\in \gamma(p)\}$.
%Calibration is the slightly stronger condition that $0$ is not the infimum of the set.
\begin{fact}[\bo{cite who?}]
  Let $\ell: \R \to \reals_+^{\Y}$ be a target loss with finite $\R$ and $\Y$.
  If a surrogate and link $(L,\psi)$ are consistent for $\ell$, then $(L,\psi)$ are calibrated for $L$.
  If $(L,\psi)$ are calibrated for $\ell$, then they indirectly elicit $\ell$.
\end{fact}

@@

\section{Upper bound: polyhedral surrogates}
In this section, we develop the proof of Theorem \ref{thm:main-upper}.
At a high level, the proof has two parts:
\begin{enumerate}
  \item Fix $p \in \simplex$.
        Prove that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for some constant $\alpha_p$.
  \item Apply this result to a finite subset $P$ of $\simplex$, namely the vertices of the level sets of $\gamma$.
        Argue (carefully) that, because a general $p$ is a convex combination of vertices of its own level set, $R_{\ell}(\psi(u),p) \leq \left(\max_{p' \in P} \alpha_{p'}\right) R_L(u,p)$.
        This implies a linear conditional transfer rate with constant $\max_{p' \in P} \alpha_{p'}$; it immediately implies a linear transfer rate as promised.
\end{enumerate}
\bo{Not sure what to say here.}
We view both parts as relatively surprising.
The first contains several important technical steps, as we will outline next.
Intuitively, it is not clear why the first part should imply the main result: it might be that $\sup_p \alpha_p = \infty$, for example, by approaching a boundary.

>>>>

\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
\usepackage{bbm}
\usepackage{amsmath,amsthm,amsfonts,amssymb}

@@

\paragraph{Target problems.}
We consider supervised machine learning problems where data points have features in $\X$ and labels in $\Y$, a finite set.
The \emph{target problem} is to learn a hypothesis mapping $\X$ to a \emph{report} (or \emph{prediction}) \emph{space} $\R$, possibly different from $\Y$.
In this paper, we assume $\R$ is a finite set.
%For example, standard binary classification has $\Y = \R = \{-1,+1\}$, whereas in a ranking problem, $\R$ is the set of all permutations of $\Y$.
We suppose the target problem is specified by the \emph{target loss} $\ell:\R\to\reals^\Y_+$, which maps a report (prediction) $r \in \R$ to the vector $\ell(r)$ of nonnegative loss values for each label, i.e. $\ell(r) = (\ell(r)_y)_{y\in\Y}$.
Given a hypothesis $g: \X \to \R$ and a data point $(x,y) \in \X \times \Y$, the loss is $\ell(g(x))_y$.
\bo{We will make a minimal non-redundancy assumption on $\ell$ in [Ref].}
We may refer to $\ell$ as a \emph{discrete} target loss to emphasize our assumption that $\R$ and $\Y$ are finite.

@@

The following definition applies both to target losses and surrogate losses.
\begin{definition}[Elicits, report space]
  \label{def:elicits}
  A loss function $f:R\to\reals^\Y_+$, where the domain $R$ is called the \emph{report space}, \emph{elicits} a property $\Gamma:\simplex \toto \R$ if
  As each $f$ elicits a unique property, we refer to it as $\prop{f}$.
\end{definition}
In particular, we will consistently use the notation $\Gamma = \prop{L}$ where $L$ is a surrogate loss with report space $\reals^d$, and we will use $\gamma = \prop{\ell}$ where $\ell$ is a target loss with finite report space $\R$.

@@

\paragraph{Calibration and consistency.}
It is known that consistency implies the weaker condition of \emph{calibration}, which implies the still-weaker, but very useful, condition of \emph{indirect elicitation}.
To define these conditions, consider $L,\psi,\ell$ and $\gamma = \prop{\ell}$.
Recall that $\gamma(p)$ of target reports that are optimal for $p$.
If $\psi(u) \not\in \gamma(p)$, then $u$ is a ``bad'' (suboptimal) surrogate prediction.
Let $B_{L,\psi,\ell}(p) = \{R_L(u,p) ~:~ \psi(u) \not\in \gamma(p)\}$, the set of expected surrogate losses resulting from bad predictions.
\begin{definition}[Indirectly elicits, Calibrated]
  A surrogate-link pair $(L,\psi)$ \emph{indirectly elicits} a target $\ell$ if $0 \not\in B_{L,\psi,\ell}(p)$ for all $p \in \simplex$.
  Meanwhile, $(L,\psi)$ is \emph{calibrated} for $\ell$ if $0 < \inf B_{L,\psi,\ell}(p)$ for all $p \in \simplex$.
%  \begin{equation}
%    \inf_{u : \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)} > \risk{L}(p)$ .
%  \end{equation}
%  Throughout, $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
\end{definition}
Equivalently, $(L,\psi)$ indirectly elicits $\ell$ if $u \in \Gamma(p) \implies \psi(u) \in \gamma(p)$.
Calibration is the slightly more subtle condition that there is a nonzero gap between the expected loss of a ``good'' report and that of any bad one.
%Observe that indirect elicitation is the condition that $0 \not\in \{\inprod{p}{L(u)} ~:~ \psi(u) \not\in \gamma(p)\}$.
%Calibration is the slightly stronger condition that $0$ is not the infimum of the set.
\begin{fact}[\bo{cite who?}]
  \label{fact:consistent-calibrated-elicits}
  Let $\ell: \R \to \reals_+^{\Y}$ be a target loss with finite $\R$ and $\Y$.
  If a surrogate and link $(L,\psi)$ are consistent for $\ell$, then $(L,\psi)$ are calibrated for $L$.
  If $(L,\psi)$ are calibrated for $\ell$, then they indirectly elicit $\ell$.
\end{fact}

@@

\section{Upper bound: polyhedral surrogates}
In this section, we develop the proof of Theorem \ref{thm:main-upper}.
At a high level, the proof has two parts:
\begin{enumerate}
  \item Fix $p \in \simplex$.
        Prove that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for some constant $\alpha_p$.
  \item Apply this result to a finite subset $P$ of $\simplex$, namely the vertices of the level sets of $\Gamma$.
        Argue (carefully) that, because a general $p$ is a convex combination of vertices of its own level set, $R_{\ell}(\psi(u),p) \leq \left(\max_{p' \in P} \alpha_{p'}\right) R_L(u,p)$.
        This implies a linear conditional transfer rate with constant $\max_{p' \in P} \alpha_{p'}$.
\end{enumerate}
We view both parts as relatively surprising.
The first part will actually hold for any calibrated surrogate.
The second part relies on a crucial fact about polyhedral losses $L$: their properties $\Gamma$ have only a finite number of full-dimensional level sets.
In other words, although the prediction space $\reals^d$ is infinite, one can pick out just a finite set of prediction points that can act as minimizers of expected loss.

\paragraph{Part 1.}
\begin{lemma} \label{lemma:fixed-p}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss and suppose the surrogate $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ are calibrated for $\ell$.
  Then for any $p \in \simplex$, there exists $\alpha_p \geq 0$ such that, for all $u \in \reals^d$,
    \[ R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p) . \]
\end{lemma}
\begin{proof}
  Fix $p \in \simplex$.
  Let $C_p = \max_{r \in \R} R_{\ell}(r,p)$.
  The maximum exists because $\ell$ is discrete, i.e. $\R$ is finite.
  Meanwhile, recall that, when defining calibration, we let $B_{L,\psi,\ell}(p) = \{R_L(u,p) ~:~ \psi(u) \not\in \gamma(p)\}$.
  Let $B_p = \inf B_{L,\psi,\ell}(p)$.
  By definition of calibration, we have $B_p > 0$.

  To combine these bounds, let $\alpha_p = \frac{C_p}{B_p}$.
  Let $u \in \reals^d$.
  There are two cases.
  If $\psi(u) \in \gamma(p)$, then $R_{\ell}(\psi(u),p) = 0 \leq R_L(u,p)$ immediately.
  If $\psi(u) \not\in \gamma(p)$, then
  \begin{align*}
    R_{\ell}(\psi(u),p)
    &\leq C_p \\
    &=    \alpha_p \cdot B_p  \\
    &\leq \alpha_p R_L(u,p) .
  \end{align*}
\end{proof}
While the above proof has the advantage of being quite short, it is also somewhat nonconstructive and potentially loose.
In \bo{Section TODO}, we investigate the constant $\alpha_p$ for polyhedral losses in particular.

\paragraph{Part 2.}
Given Lemma \ref{lemma:fixed-p}, we might hope to obtain that, for any calibrated surrogate, there exists $\alpha = \sup_p \alpha_p$ such that $R_{\ell}(\psi(u),p') \leq \alpha R_L(u,p')$ for all $p'$.
However, we know this is false in general: not all surrogates yield linear surrogate regret transfer rates!
Indeed, for many surrogates, the supremum will be $+\infty$.

However, polyhedral surrogates have a special structure that allows us to show $\alpha$ is finite.
We first present several general lemmas, then the key implication from $L$ being polyhedral.
The general results are known \bo{cite}.
\begin{lemma} \label{lemma:refines}
  If $(L,\psi)$ indirectly elicits $\ell$, then $\Gamma = \prop{L}$ \emph{refines} $\gamma = \prop{\ell}$ in the sense that, for all $u \in \reals^d$, there exists $r \in \R$ such that $\Gamma_u \subseteq \Gamma_r$.
\end{lemma}
\begin{proof}
  For any $u$, let $r = \psi(u)$.
  By indirect elicitation, $u \in \Gamma(p) \implies r \in \gamma(p)$.
  Phrased differently, $\Gamma_u = \{p ~:~ u \in \Gamma(p)\} \subseteq \{p ~:~ r \in \gamma(p)\} = \gamma_r$.
\end{proof}
The next lemma states that surrogate regret is linear in $p$ on any fixed level set $\Gamma_{u^*}$ of $\Gamma = \prop{L}$.
By combining this fact with Lemma \ref{lemma:refines}, we obtain that target regret is linear on these level sets as well.
\begin{lemma} \label{lemma:linear-on-levelset}
  Suppose $(L,\psi)$ indirectly elicits $\ell$ and let $\Gamma = \prop{L}$.
  Then for any fixed $u,u^* \in \reals^d$ and $r \in \R$, the functions $R_L(u,\cdot)$ and $R_{\ell}(u,\cdot)$ are linear in their second arguments on $\Gamma_{u^*}$.
\end{lemma}
\begin{proof}
  Let, $p \in \Gamma_{u^*}$.
  By definition on $\Gamma_{u^*}$, $\risk{L}(p) = \inprod{p}{L(u^*)}$.
  So for fixed $u$,
    \[ R_L(u,p) = \inprod{p}{L(u)} - \inprod{p}{L(u^*)} = \inprod{p}{L(u) - L(u^*)} , \]
  a linear function of $p$.
  Next, by Lemma \ref{lemma:refines}, there exists $r^*$ such that $\Gamma_{u^*} \subseteq \gamma_{r^*}$.
  By the same argument, for fixed $r$, $R_{\ell}(r,p) = \inprod{p}{\ell(r) - \ell(r^*)}$, a linear function.
\end{proof}

We now arrive at the key property of polyhedral losses: \emph{although the prediction space is $\reals^d$, we can identify a finite set of predictions that characterize $L$.} \bo{wording?}
Here, a level set is \emph{full-dimensional} if it has affine dimension $|\Y|-1$, the same as the simplex $\simplex$.
\begin{lemma} \label{lemma:polyhedral-finite}
  If $L: \reals^d \to \reals_+^{\Y}$ is polyhedral, then $\Gamma = \prop{L}$ has a finite set of level sets that union to $\simplex$. Also, these level sets are polytopes.
\end{lemma}
\begin{proof}
  \bo{prove, cite, etc}
\end{proof}

We are now ready to restate and prove the main upper bound.
\begin{theorem}[Theorem \ref{thm:main-upper} restated]
  Let the surrogate loss $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ be consistent for the target loss $\ell: \R \to \reals_+^{\Y}$.
  If $L$ is polyhedral, then $(L,\psi)$ guarantee a linear regret transfer rate for $\ell$, i.e. there exists $\alpha > 0$ such that, for all $\D$ and all measurable $h: \X \to \reals^d$,
    \[ R_{\ell}(\psi \circ h ; \D) \leq \alpha R_L(h ; \D) . \]
\end{theorem}
\begin{proof}
  We first recall that by Fact \ref{fact:consistent-calibrated-elicits}, consistency implies that $(L,\psi)$ are calibrated for $\ell$ and that $(L,\psi)$ indirectly elicit $\ell$.
  Next, by Observation \ref{obs:transfer}, it suffices to show a linear \emph{conditional} regret transfer rate, i.e. for all $p \in \simplex$ and $u \in \reals^d$, we show $R_{\ell}(\psi(u),p) \leq \alpha R_L(u,p)$.

  By Lemma \ref{lemma:polyhedral-finite}, the polyhedral loss $L$ has a finite set $U \subset \reals^d$ of predictions such that (a) for each $u \in U$, the level set $\Gamma_u$ is a polytope, and (b) $\cup_{u \in U} \Gamma_u = \simplex$.
  Let $\mathcal{Q}_u \subseteq \simplex$ be the finite set of vertices of the polytope $\Gamma_u$, and let $\mathcal{Q} = \cup_{u \in U} \mathcal{Q}_u$.

  By Lemma \ref{lemma:fixed-p}, for each $p \in \mathcal{Q}$, there exists $\alpha_p \geq 0$ such that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for all $u$.
  We choose
    \[ \alpha = \max_{p \in \mathcal{Q}} \alpha_p . \]
  To prove the conditional transfer rate, consider any $p \in \simplex$ and any $u \in \reals^d$.
  There exists $u \in U$ such that $p \in \Gamma_u$, so we can write $p$ as a convex combination of the vertices, i.e.
    \[ p = \sum_{p' \in \mathcal{Q}_u} \beta(p') p' \]
  for some probability distribution $\beta$.
  Because $\mathcal{Q}_u \subseteq \Gamma_u$, $R_L$ and $R_{\ell}$ are linear on it by Lemma \ref{lemma:linear-on-levelset}.
  So:
  \begin{align*}
    R_{\ell}(u,p)
    &=    R_{\ell}\left(\psi(u) ~,~ \sum_{p' \in \mathcal{Q}_u} \beta(p') p'\right)  \\
    &=    \sum_{p' \in \mathcal{Q}_u} \beta(p') R_{\ell}(\psi(u),p')  \\
    &\leq \sum_{p' \in \mathcal{Q}_u} \beta(p') \alpha_{p'} R_L(u,p')  \\
    &\leq \alpha \sum_{p' \in \mathcal{Q}_u} \beta(p') R_L(u,p')  \\
    &=    \alpha R_L(u, p) .
  \end{align*}
\end{proof}
